Guarded Recursion in Agda via Sized Types

### 2019-02-01 code-oneclock-paper/latex/paper.tex
abstract

\begin{abstract}
Coinduction allows programming and reasoning with possibly nonterminating programs and potentially infinite objects.
In this paper, we compare two approaches for extending type theory with coinduction and coinductive types.
More specifically, we show that guarded recursion and guarded recursive types can be simulated with sized types.
We do this by defining a model of guarded recursion in type theory extended with sized types.
The model is variation of the topos of trees, and it is formalized in Agda.
\end{abstract}

@@

This work can be extended is several different directions.
Various type theories for guarded recursion in the literature include
dependent types. Currently, there exist two disciplines for mixing
dependent types with guarded recursion: delayed substitutions
\cite{bizjak2016guarded} and ticks \cite{BahrGM17}. It would be interesting to extend
the syntax and semantics of \GTT\ with dependent types following one
of these methods.

>>>>

\begin{abstract}
  In type theory, programming and reasoning with possibly
  non-terminating programs and potentially infinite objects is
  achieved using coinductive types. Crucially, recursively defined
  programs of coinductive types are required to be productive. Proof
  assistants such as Agda and Coq traditionally employ strict syntactic
  productivity checks, which commonly make programming with coinductive types
  convoluted. One way to overcome this issue is by encoding productivity at
  the level of types, so that the type system forbids the
  implementation of non-productive corecursive programs.
%%, meaning
%%  that although their definitions can be unfolded \emph{ad infinitum},
%%  finite unfoldings are guaranteed to produce . 
  In this paper we compare two different approaches to type-based
  productivity: guarded recursion and sized types. More specifically,
  we show how to simulate guarded recursion in Agda using sized
  types. First we formalize the syntax of a simple type theory for
  guarded recursion, corresponding to a variant of Atkey and McBride's
  type system for productive coprogramming
  \cite{atkey2013productive}. Then we give a denotational semantics
  using presheaves over the preorder of sizes. Sized types are
  fundamentally employed in the interpretation of the characteristic
  features of guarded recursion, notably the fixpoint combinator.
%%extending type theory with
%%  coinduction and coinductive types.  More specifically, we show that
%%  guarded recursion and guarded recursive types can be simulated with
%%  sized types.  We do this by defining a model of guarded recursion in
%%  type theory extended with sized types.  The model is variation of
%%  the topos of trees, and it is formalized in Agda.
\end{abstract}

@@

This work can be extended in several different directions.
Various type theories for guarded recursion in the literature include
dependent types. Currently, there exist two disciplines for mixing
dependent types with guarded recursion: delayed substitutions
\cite{bizjak2016guarded} and ticks \cite{BahrGM17}. It would be interesting to extend
the syntax and semantics of \GTT\ with dependent types following one
of these methods.
